\section{Crys2Seward}

Crys2seward is a python program that links the output from the CRYSTAL geometry optimisation to \texttt{disp\_solve}, \texttt{xenv15} and \texttt{env2seward} to produce a grid of input files for the program SEWARD.

\subsection{Installing python 3}

A \textbf{python 3} distribution must be installed. Either Anaconda or direct installation \url{https://realpython.com/installing-python/} are arguably the best options. 

\subsection{Correct packages}

The following packages must be installed with your distribution for the program to run:

\begin{itemize}
	\item sys
	\item os
	\item shutil
	\item pathlib
	\item env2seward
	\item disp\_solve
\end{itemize}

Install packages you don't have with \texttt{pip install <module\_name>} or check that you have it with \texttt{pip list | grep <module\_name\_you\_want\_to\_check>}, unless they are packages available only in this repository such as env2seward and disp\_solve.

In that case you will need to download both the script and the \texttt{\_\_init\_\_.py} files from the git repository, and reference their path in \texttt{crys2seward.py} when declaring used packages (unless they are located in the same relative path as the repository, as it is by default, then no changes are required).

\subsection{Input files}

The program requires the following inputs:

\begin{itemize}
	
	\item CRYSTAL output file of a phonon calculation of your system including IR intensities (Born charges) and Hessian matrix printed in fractional coordinates
	\item The initial system.cell file of the ENV code corresponding to the undistorted system
	\item The envin file for the env15 program \texttt{xenv15} after executing the necessary \texttt{make} command in the Env15 folder (see the local tools manual written by M. B. Lepetit for how to install Env15).
	\item The prefix.c2s.in input file containing all the paths pointing to the aforementioned files, as well as the inputs for \texttt{disp\_solve} and \texttt{env2seward}
	
	\item (Optional) The .DAT files for the Born and Hessian matrices, used if the CRYSTAL output .loto.out file does not contain the necessary data to construct one or both of these matrices.
\end{itemize}


\subsection{Executing the script}

To execute the script, type:

\texttt{python3 <path\_to\_crys2seward.py> <path\_to\_c2s.in\_input\_file>}

into the command line in the same working directory as the xenv15 program and it's input files (envin and env.out).
\texttt{xenv15} only works in the current working directory, i.e. it cannot be referenced from another path.

For example, on this git repository the relative paths are:

\texttt{python3 ../../../python\_scripts/crys2seward/crys2seward.py ../ymno3\_d1.c2s.in}

The main difficulty of executing the script is writing a correct input file. See figure (6) for an example.

\begin{figure}
	\centering
	\begin{BVerbatim}[fontsize=\small]
	#########crystal2seward input########################
	envin = ymno3_d1.envin
	envout = ymno3_d1.env.out
	xenv_sew0 = ymno3_d1.sew0
	xenv_psd = ymno3_d1.psd
	#########disp_solve input############################
	born_file = ../ex_student_ymno3_data/Position/YMnO3/BORN_B1Pw_loto.DAT
	hess_file = ../ex_student_ymno3_data/Position/YMnO3/HESSIEN.DAT
	crystal_file = ../ex_student_ymno3_data/frequence.B1PW.loto.out
	cell_init = ../ymno3.cell
	ex = [0,.15,.03]
	ey = [0,.15,.03]
	ez = [0,.15,.03]
	unit_source = auto
	charge_dict = {'Y':3.0, 'MN':3.0, 'O1':-2.0, 'O2':-2.0}
	##########env2seward input###########################
	filename = ymno3_d1.sew.in
	title = <insert_title_here>
	sew0_file = ymno3_d1.0.0_0.0_0.0.env.sew0
	psd_file = ymno3_d1.0.0_0.0_0.0.env.psd
	lib_frag = {'O': {'key': 'O_frag_basis', 'loc': ''},
	'Y': {'key': 'Y_frag_basis', 'loc': ''}, 'MN': {'key': 'MN_frag_basis', 'loc': ''}}
	lib_pseudo = {'MN': {'key': 'MN_pseudo_basis', 'loc': 'MN_pseudo_library'},
	'O1': {'key': 'O1_pseudo_basis', 'loc': 'O1_pseudo_library'},
	'O2': {'key': 'O2_pseudo_basis', 'loc': 'O2_pseudo_library'},
	'O3': {'key': 'O3_pseudo_basis', 'loc': 'O3_pseudo_library'},
	'O4': {'key': 'O4_pseudo_basis', 'loc': 'O4_pseudo_library'},
	'Y': {'key': 'Y_pseudo_basis', 'loc': 'Y_pseudo_library'},
	'Y1': {'key': 'Y1_pseudo_basis', 'loc': 'Y1_pseudo_library'},
	'Y2': {'key': 'Y2_pseudo_basis', 'loc': 'Y2_pseudo_library'}}
	\end{BVerbatim}
	\caption{}
\end{figure}

\subsubsection{Rules for c2s.in:}

\begin{itemize}
	\item envin and envout correspond to the absolute path of the prefix.envin, prefix.env.out files
	\item xenv\_sew0 and xenv\_psd correspond to the default names of the xenv15 output files for the system
	\item The disp\_solve and env2seward input rules can be found in their README's and their sections in this software manual.
\end{itemize}

\subsection{Output}

Each grid of files is stored in an appropriately named folder. The sew0, psd and sew.in files are generated in separate folders but in the same current working directory with the xenv15 files as discussed before.
The disp\_solve cell files are generated in the same way, but stored in the same directory as the initial cell file or the CRYSTAL output file if their relative or full paths are specified.
They are \textbf{not} generated in the local directory unless the disp\_solve input file paths are not specified. Every grid folder name features the start, stop, step for Ex, Ey, Ez chosen in the c2s.in file in the disp\_solve input section.